---
title : "Test of the migration of temperate species at their northern limit 
with a glmm model on space and time"
author : "Clémentine de Montgolfier"
date : "2023-06-23"
output : pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "../../")
```

```{r library, include = FALSE}
library(tidyverse)
library(glmmTMB)
library(patchwork)
library(DHARMa)
library(reshape2)
```

# Introduction et question

Study of the presence and absence of my temperate species on the bioclimatic domains
of the yellow and white birch fir stands.

For now with the data of the gaules of all size classes.

Problem : the graph of the abundances does not take into account the presence,
maybe we should make a graph with the two combined (see § combination of presence and abundance)

# Data

Data import and selection :
- climatic domain selection : 4 (sapinière à bouleau jaune) and 5 (sapinière à bouleau blanc)
- species selection : ACERUB (Acer rubrum), ACESAC (Acer saccharum), BETALL (betula alleghaniensis)

```{r data}
load("data/full_data.RData")
```

# Model with glmmTMB

Model de hurdle :
- 1st part : logistic regression on the presence/absence of the species (bernouilli distribution)
- 2nd part : linear regression on the abundance of the species (truncated poisson distribution)

With :
- individual run for each species
- random effect of the plot : id_pe
- fixed effect and interaction of the year and the latitude

```{r}
model_sp <- function(sp, part){
    if(part == "presence"){
    return(glmmTMB(
        presence_gaule ~ latitude_sc*year_measured_sc +
            (1|id_pe),
        data = full_data %>% filter(sp_code == sp),
        family = binomial(link = "logit")))
    }
    if(part == "abundance"){
    return(glmmTMB(
    all_cl ~ latitude_sc*year_measured_sc +
            (1|id_pe),
    data = full_data %>% filter(sp_code == sp, presence_gaule == 1),
    family = truncated_poisson(link = "log")))
    }
}
```

```{r}
Mysp = c("ACERUB", "ACESAC", "BETALL")
for (s in Mysp){
    for(p in c("presence", "abundance")){
        assign(paste0(s, "_", p), model_sp(s,p))
    }
}
```

# Model check

Model check :
- residual distribution
- residual vs fitted values
- over/underdispersion

```{r model check}
list_model = list(ACESAC_presence, ACERUB_abundance, ACESAC_presence, ACESAC_abundance, BETALL_presence, BETALL_abundance)
for (model in list_model) {
    resid_sim <- simulateResiduals(model, integerResponse = TRUE)
    plot(resid_sim)
}
```

# Predictions

```{r}
time_grid <- seq(
    min(full_data$year_measured_sc),
    max(full_data$year_measured_sc),
    length.out = 50) #mh
# Créer une grille pour la latitude
latitude_grid <- seq(
    quantile(full_data$latitude_sc,.05), #mh
    quantile(full_data$latitude_sc,.95), #mh
    length.out = 50) #mh
new_data <- expand.grid(
            year_measured_sc = time_grid,
            latitude_sc = latitude_grid,
            latitude2_sc = latitude_grid^2,
            id_pe = 1)
# faire les colonnes latitudes et year_measured unscalled
mean_year = scaling[which(scaling$var == "year_measured" & scaling$mean_sd == "mean"),]$value %>% as.numeric()
sd_year = scaling[which(scaling$var == "year_measured" & scaling$mean_sd == "sd"),]$value %>% as.numeric()
mean_lat = scaling[which(scaling$var == "latitude" & scaling$mean_sd == "mean"),]$value %>% as.numeric()
sd_lat = scaling[which(scaling$var == "latitude" & scaling$mean_sd == "sd"),]$value %>% as.numeric()
new_data$latitude <- new_data$latitude_sc * sd_lat + mean_lat
new_data$year_measured <- new_data$year_measured_sc * sd_year + mean_year

new_data$pred_ACERUB_presence <- predict(ACERUB_presence, newdata = new_data, type = "response", allow.new.levels = TRUE)
new_data$pred_ACERUB_abundance <- predict(ACERUB_abundance, newdata = new_data, type = "response", allow.new.levels = TRUE)
new_data$pred_ACERUB = new_data$pred_ACERUB_presence * new_data$pred_ACERUB_abundance
new_data$pred_ACESAC_presence <- predict(ACESAC_presence, newdata = new_data, type = "response", allow.new.levels = TRUE)
new_data$pred_ACESAC_abundance <- predict(ACESAC_abundance, newdata = new_data, type = "response", allow.new.levels = TRUE)
new_data$pred_ACESAC = new_data$pred_ACESAC_presence * new_data$pred_ACESAC_abundance
new_data$pred_BETALL_presence <- predict(BETALL_presence, newdata = new_data, type = "response", allow.new.levels = TRUE)
new_data$pred_BETALL_abundance <- predict(BETALL_abundance, newdata = new_data, type = "response", allow.new.levels = TRUE)
new_data$pred_BETALL = new_data$pred_BETALL_presence * new_data$pred_BETALL_abundance
# remplacer les 0 par NA
#new_data$pred_ACERUB[new_data$pred_ACERUB == 0] <- NA

#wd <- new_data %>% pivot_longer(cols = c("pred_ACERUB_abundance", "pred_ACESAC_abundance", "pred_BETALL_abundance"), names_to = "sp_code", values_to = "pred")
wd <- new_data %>% pivot_longer(cols = c("pred_ACERUB_presence", "pred_ACESAC_presence", "pred_BETALL_presence"), names_to = "sp_code", values_to = "pred")
#wd <- new_data %>% pivot_longer(cols = c("pred_ACERUB", "pred_ACESAC", "pred_BETALL"), names_to = "sp_code", values_to = "pred")
# colonne sp code : selectionner seulement les 6 derniers caractères
wd$sp_code <- substr(wd$sp_code, 6, 11)
```

# Figures

Plot of the predictions :
- for each species
- for each part of the model (presence/absence and abundance)
- with : 
    - the latitude on the x axis
    - the year on the y axiswd

    - the color for the prediction

last paragraphe in english :

Table for all species and the two parts of the model :
- with the presence probabilities
- with the abundances
- with a column for the species (sp_code)

```{r}
library("RColorBrewer")
myPalette1 <- colorRampPalette(c("#1ee4e7", "#1ee4e7", "#17b91f", "#fffb00", "#e5a836", "#e26e10", "#c00606", "#670101"))
# créer ma propre palette de couleur avec 20 couleurs et des plus grands écart de couleurs au début
myPalette2 <- colorRampPalette(c("white","#001b92","#1ee4e7", "#14af1c", "#fffb00", "#e5a836", "#e26e10", "#c00606", "#670101"))
myPalette3 <- colorRampPalette(c("white","#1ee4e7", "#fffb00","#e5a836", "#c00606"))
myPalmh <- colorRampPalette(c("white", rev(brewer.pal(10, "Spectral")))) #mh
```

```{r combination prediction plot}
combination_plot <-
    ggplot(wd) +
        aes(y = latitude, x = year_measured, fill = (pred)) +
        geom_raster(interpolate = TRUE) +
        scale_fill_gradientn(colours = myPalmh(200), labels = function(x) format(x, scientific = TRUE)) +
        facet_wrap(~sp_code, labeller = as_labeller(c("ACERUB" = "Acer rubrum", "ACESAC" = "Acer saccharum",
                        "BETALL" = "Betula alleghaniensis"))) +
        scale_x_continuous(expand=c(0,0)) +
        scale_y_continuous(expand=c(0,0)) +
        # Ecrire les facettes en haut sans background
        theme(strip.background = element_blank(),
              strip.text.x = element_text(size = 14, face = "italic", color = "black", hjust = 0),
              strip.placement = "outside") +
        # afficher les lignes des axes sur toutes les facettes
        theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) +
    # limite entre les facettes
        theme(panel.spacing = unit(2, "lines")) +
        # gradiant legend : plot height
        guides(fill = guide_colourbar(barheight = 15)) +
        labs(y = "Latitude", x = "Années", fill = "P(présence)")
```

```{r print combination plot}
combination_plot
```

# Pour les autres espèces

```{r}
Mysp = c("ABIBAL","PICMAR","PICGLA","BETPAP","POPTRE")
for (s in Mysp){
    for(p in c("presence", "abundance")){
        assign(paste0(s, "_", p), model_sp(s,p))
    }
}
```

```{r}
time_grid <- seq(
    min(full_data$year_measured_sc),
    max(full_data$year_measured_sc),
    length.out = 50) #mh
# Créer une grille pour la latitude
latitude_grid <- seq(
    quantile(full_data$latitude_sc,.05), #mh
    quantile(full_data$latitude_sc,.95), #mh
    length.out = 50) #mh
new_data <- expand.grid(
            year_measured_sc = time_grid,
            latitude_sc = latitude_grid,
            latitude2_sc = latitude_grid^2,
            id_pe = 1)
# faire les colonnes latitudes et year_measured unscalled
mean_year = scaling[which(scaling$var == "year_measured" & scaling$mean_sd == "mean"),]$value %>% as.numeric()
sd_year = scaling[which(scaling$var == "year_measured" & scaling$mean_sd == "sd"),]$value %>% as.numeric()
mean_lat = scaling[which(scaling$var == "latitude" & scaling$mean_sd == "mean"),]$value %>% as.numeric()
sd_lat = scaling[which(scaling$var == "latitude" & scaling$mean_sd == "sd"),]$value %>% as.numeric()
new_data$latitude <- new_data$latitude_sc * sd_lat + mean_lat
new_data$year_measured <- new_data$year_measured_sc * sd_year + mean_year


for (sp in Mysp){
    new_data <- new_data %>% mutate(!!paste0("pred_", sp, "_presence") := predict(get(paste0(sp, "_presence")), newdata = new_data, type = "response", allow.new.levels = TRUE),
                                    !!paste0("pred_", sp, "_abundance") := predict(get(paste0(sp, "_abundance")), newdata = new_data, type = "response", allow.new.levels = TRUE),
                                    !!paste0("pred_", sp) := !!sym(paste0("pred_", sp, "_presence")) * !!sym(paste0("pred_", sp, "_abundance")))
}

wd <- new_data %>% pivot_longer(cols = c("pred_ABIBAL", "pred_PICMAR", "pred_PICGLA", "pred_BETPAP", "pred_POPTRE"), names_to = "sp_code", values_to = "pred")
# colonne sp code : selectionner seulement les 6 derniers caractères
wd$sp_code <- substr(wd$sp_code, 6, 11)
```

```{r combination prediction plot reste}
for (sp in Mysp){
    assign(paste0("combination_plot_", sp),
        ggplot(wd %>% filter(sp_code == sp)) +
        aes(y = latitude, x = year_measured, fill = (pred)) +
        geom_raster(interpolate = TRUE) +
        scale_fill_gradientn(colours = myPalette3(20), labels = function(x) format(x, scientific = TRUE)) +
        facet_wrap(~sp_code, labeller = as_labeller(c("ABIBAL" = "A. balsamea", "PICMAR" = "P. mariana",
                        "PICGLA" = "P. glauca", "BETPAP" = "B. papyrifera", "POPTRE" = "P. tremuloides"))) +
        scale_x_continuous(expand=c(0,0)) +
        scale_y_continuous(expand=c(0,0)) +
        # Ecrire les facettes en haut sans background
        theme(strip.background = element_blank(),
              strip.text.x = element_text(size = 10, face = "italic", color = "black", hjust = 0),
              strip.placement = "outside") +
        # afficher les lignes des axes sur toutes les facettes
        theme(panel.border = element_rect(colour = "black", fill=NA, size=0.5)) +
    # limite entre les facettes
        theme(panel.spacing = unit(2, "lines")) +
        # gradiant legend : plot height
        guides(fill = guide_colourbar(barheight = 20)) +
        labs(y = "Latitude", x = "Year", fill = "Sapling number"))
}
```

```{r}

combination_plot_ABIBAL + 
combination_plot_PICMAR + 
combination_plot_PICGLA +
combination_plot_BETPAP + 
combination_plot_POPTRE
```