---
output : html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE)
# directory one up
knitr::opts_knit$set(root.dir = "../../")
```

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(patchwork)
library(glmmTMB)
library(buildmer)
```

# 1. Load data and models

Load the data

```{r data, echo = FALSE}
# load data
Mysp = c("ABIBAL","ACERUB", "ACESAC", "BETALL","PICMAR", "PICGLA", "BETPAP", "POPTRE")
load("data/full_data.RData")
# This data countains full_data a dataframe with the data I analysed
# and scaling a dataframe with the scaling factor for each variables (mean and sd)
full_data <- full_data %>% 
    mutate(n_perturb = is_logging + is_partial_logging + is_logging_pr + is_burn + is_outbreak) %>%
    filter(dom_bio %in% c(4,5), n_perturb == 0)
```

```{r}
full_data$epmatorg_sc2 <- full_data$epmatorg_sc^2
full_data$ph_humus_sc2 <- full_data$ph_humus_sc^2
full_data$cmi_sum_sc2 <- full_data$cmi_sum_sc^2
full_data$an_meanT_sc2 <- full_data$an_meanT_sc^2
full_data$tree_nb_sp2 <- full_data$tree_nb_sp^2
full_data <- full_data %>% group_by(id_pe) %>% mutate(prev_gaule = lag(all_cl))
```

# 2. Models

Model : buldglmmTMB -> make the model and perform the step selection

```{r, message = FALSE, warning = FALSE, eval = FALSE}
model_sp <- function(sp, part){
    if(part == "presence"){
    return(buildglmmTMB(
        presence_gaule ~ epmatorg_sc + epmatorg_sc2 + ph_humus_sc + ph_humus_sc2 +
            an_meanT_sc + an_meanT_sc2 + tree_ba_sc +
            tree_nb_sp + tree_nb_sp2 + cmi_sum_sc + cmi_sum_sc2 + texture +
            (1|id_pe),
        data = full_data %>% filter(sp_code == sp),
        family = binomial(link = "logit")))
    }
    if(part == "abundance"){
    return(buildglmmTMB(
    all_cl ~ epmatorg_sc + epmatorg_sc2 + ph_humus_sc + ph_humus_sc2 +
        an_meanT_sc + an_meanT_sc2 + tree_ba_sc +
        tree_nb_sp + tree_nb_sp2 + cmi_sum_sc + cmi_sum_sc2 + texture +
        (1|id_pe),
    data = full_data %>% filter(sp_code == sp, presence_gaule == 1),
    family = truncated_poisson(link = "log")))
    }
}
```

```{r, message = FALSE, warning = FALSE, eval = FALSE}
for (s in Mysp){
    for(p in c("presence", "abundance")){
        assign(paste0(s, "_", p), model_sp(s,p))
    }
}
```

# 3. Results

```{r, eval = FALSE}
result <- data.frame()

for(i in 1:5){
    for (sp in Mysp){
        for(p in c("presence", "abundance")){
            df <- data.frame(AIC = AIC(get(paste0(sp, "_", p, i))), sp_code = sp, part = p, model = i)
            result <- rbind(result, df)
        }
    }
}
saveRDS(result, "result.RDS")
```

```{r, eval = FALSE}
result <- readRDS("result.RDS")
# pivot wider on models with AIC in columns named (AIC1, AIC2, AIC3, AIC4)
result_wide <- result %>% pivot_wider(names_from = model, values_from = AIC)
```

# 4. Selection process

Run once with :
        presence_gaule/all_cl ~ epmatorg_sc + epmatorg_sc2 + ph_humus_sc + ph_humus_sc2 + altitude_sc + texture +
            cmi_sum_sc + cmi_sum_sc2 + an_meanT_sc + an_meanT_sc2 + tree_ba_sc +
            tree_nb_sp + tree_nb_sp2 +
            (1|id_pe)

Then look at the number of time each variables are significant

Removed (less significant or correlated) :
- Texture
- altitude ~ Temperature + cmi
- cmi et cmi2

Run everything again

Confirm the drop of selected variables by comparing all AIC

### Model1 : TOUT

| presence_gaule/all_cl ~ 
|     epmatorg_sc + epmatorg_sc2 +
|     ph_humus_sc + ph_humus_sc2 +
|     texture +
|     cmi_sum_sc + cmi_sum_sc2 +
|     an_meanT_sc + an_meanT_sc2 +
|     altitude_sc +
|     tree_ba_sc +
|     tree_nb_sp + tree_nb_sp2 +
|     (1|id_pe)

### Model2 : -TEXTURE

| presence_gaule/all_cl ~ 
|     epmatorg_sc + epmatorg_sc2 +
|     ph_humus_sc + ph_humus_sc2 +
|     cmi_sum_sc + cmi_sum_sc2 +
|     an_meanT_sc + an_meanT_sc2 +
|     altitude_sc +
|     tree_ba_sc +
|     tree_nb_sp + tree_nb_sp2 +
|     (1|id_pe)

### Model3 : -ALTITUDE

| presence_gaule/all_cl ~ 
|     epmatorg_sc + epmatorg_sc2 + 
|     ph_humus_sc + ph_humus_sc2 +
|     cmi_sum_sc + cmi_sum_sc2 + 
|     an_meanT_sc + an_meanT_sc2 + 
|     tree_nb_sp + tree_nb_sp2 +
|     tree_ba_sc +
|     (1|id_pe)

### Model4 : -CMI ET CMI2

| presence_gaule/all_cl ~ 
|     epmatorg_sc + epmatorg_sc2 + 
|     ph_humus_sc + ph_humus_sc2 +
|     an_meanT_sc + an_meanT_sc2 + 
|     tree_nb_sp + tree_nb_sp2 +
|     tree_ba_sc +
|     (1|id_pe)

## Model5 : -ALTITUDE Mais avec texture

| presence_gaule/all_cl ~ 
|     epmatorg_sc + epmatorg_sc2 + 
|     ph_humus_sc + ph_humus_sc2 +
|     texture +
|     cmi_sum_sc + cmi_sum_sc2 + 
|     an_meanT_sc + an_meanT_sc2 + 
|     tree_nb_sp + tree_nb_sp2 +
|     tree_ba_sc +
|     (1|id_pe)


### All model AIC

```{r}
sp_code <- c("ABIBAL", "ABIBAL", "ACERUB", "ACERUB", "ACESAC", "ACESAC", "BETALL", "BETALL", "PICMAR", "PICMAR", "PICGLA", "PICGLA", "BETPAP", "BETPAP", "POPTRE", "POPTRE")
part <- c("presence", "abundance", "presence", "abundance", "presence", "abundance", "presence", "abundance", "presence", "abundance", "presence", "abundance", "presence", "abundance", "presence", "abundance")
Model_1 <- c(3464, 16898, 1740, 1506, 880, 1228, 776, 360, 1972, 7635, 1773, 1027, 3105, 4610, 868, 527)
Model_2 <- c(3471, 8861, 1740, 1506, 880, 1228, 774, 360, 1982, 7752, 1781, 740, 3105, 3272, 868, 527)
Model_3 <- c(3479, 8793, 1740, 1506, 884, 1227, 774, 361, 1986, 7785, 1781, 734, 3104, 4670, 868, 524)
Model_4 <- c(3485, 17012, 1740, 1512, 886, 1228, 797, 361, 1997, 7809, 1800, 740, 3146, 4771, 874, 524)
Model_5 <- c(3469, 8777, 1739, 1506, 869, 1227, 774, 361, 1976, 7635, 1773, 1049, 3104, 3269, 867, 524)

result_wide <- data.frame(sp_code, part, Model_1, Model_2, Model_3, Model_4, Model_5)
result_wide %>% knitr::kable()
```

## Model avec AIC le plus petit (ou < 2 de différence avec le plus petit)

Au départ je n'avais fait que les 4 premiers models

```{r}
min_result <- result_wide %>%
    pivot_longer(cols = c(Model_1, Model_2, Model_3, Model_4), names_to = "model", values_to = "AIC") %>%
    group_by(sp_code, part) %>% filter(AIC == min(AIC) | abs(AIC - min(AIC)) <= 2)
table(min_result$model) %>% knitr::kable()
```

Et j'en avais choisis le Model 3

Mais si je rajoute la texture il devisns meilleur pour plusieurs espèces

```{r}
min_result <- result_wide %>%
    pivot_longer(cols = c(Model_1, Model_2, Model_3, Model_4, Model_5), names_to = "model", values_to = "AIC") %>%
    group_by(sp_code, part) %>% filter(AIC == min(AIC) | abs(AIC - min(AIC)) <= 2)
table(min_result$model) %>% knitr::kable()
```

A voir si on ne peut pas "expliquer la texture par les autres variables", corélation, j'ai aussi 72 0 qui correspondent probablement à des NA,
Soit je les supprime, soit je trouve une manière de travailler avec (c'est possible) 

# 5. Correlation

```{r}
library(corrplot)
full_data %>% ungroup() %>% 
    select(epmatorg_sc, ph_humus_sc, cmi_sum_sc, an_meanT_sc, tree_ba_sc, tree_nb_sp, altitude_sc) %>%
    cor() %>% corrplot(method = "number")
```

```{r}
ggplot(full_data %>% filter(sp_code == "ACERUB")) +
    geom_boxplot(aes(x = texture, y = epmatorg)) +
ggplot(full_data %>% filter(sp_code == "ACERUB")) +
    geom_boxplot(aes(x = texture, y = ph_humus)) +
ggplot(full_data %>% filter(sp_code == "ACERUB")) +
    geom_boxplot(aes(x = texture, y = an_meanT)) +
ggplot(full_data %>% filter(sp_code == "ACERUB")) +
    geom_boxplot(aes(x = texture, y = cmi_sum))
table(full_data$texture ==0)
```

Altitude is correlated with temperature, cmi and ph (maybe remove)
Removed altitude.
Test avec cmi car corrélé à Temp

# 6. Choix

Keep all same variables for every model
Keep random effect even if dropped

# 7. Conclusion

Keep model 3 with :

presence_gaule/all_cl ~ epmatorg_sc + epmatorg_sc2 + ph_humus_sc + ph_humus_sc2 +
    cmi_sum_sc + cmi_sum_sc2 + an_meanT_sc + an_meanT_sc2 + tree_ba_sc +
    tree_nb_sp + tree_nb_sp2 +
    (1|id_pe)

C'est avec ça que j'ai fait le model Bayesien mais Model 5 peut-être mieux.