---
output : html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE)
# directory one up
knitr::opts_knit$set(root.dir = "../../")
```

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(patchwork)
library(R2jags)
library(boot)
library(cowplot)
```

# 1. Load data and models

Load the data

```{r data, echo = FALSE}
# load data
Mysp = c("ABIBAL","ACERUB","BETALL", "ACESAC","PICMAR", "PICGLA", "BETPAP", "POPTRE")
load("data/full_data.RData")
# This data countains full_data a dataframe with the data I analysed
# and scaling a dataframe with the scaling factor for each variables (mean and sd)
full_data <- full_data %>% filter(sp_code %in% Mysp, dom_bio %in% c(4,5))
```

Load the models output for every species

```{r}
Model_path = "output/heavy/output_new_withoutBA/" # impossible to upload on github for now because to big
for (sp in Mysp){
    assign(paste0("Model_", sp), readRDS(paste0(Model_path, sp, ".rds")))
}
```

Load the DIC and deviance for every species

```{r}
DIC <- data.frame()
for (sp in Mysp){
    model <- get(paste0("Model_", sp))
    DIC <- rbind(DIC, data.frame(DIC = model$BUGSoutput$DIC, sp_code = sp))
}

deviance <- data.frame()
for (sp in Mysp){
    model <- get(paste0("Model_", sp))
    deviance <- rbind(deviance, data.frame(deviance = model$BUGSoutput$summary["deviance","mean"], sp_code = sp))
}
```

Load the function for making data from the models and plotting

```{r}
source("R/4_Disturbance_model/function_analysis.R")
```

# 2. Model output for all coefficients except disturbance related ones

```{r, fig.height=8, fig.width=10}
# Extract the coefficient and their 95% confidence interval
all_param <- all_param_table()
# Plot the results for all coefficients
all_param_plot()
# save all_param data as rds
saveRDS(all_param, "output/all_param.rds")
```

```{r}
list_coef = 
    c("Conspecific", "Conspecific2","Basal area", "Humus pH","Organic matter", "Mean temperature", "CMI")
ggplot(all_param %>% filter (var %in% list_coef, type == "pa")) +
    geom_point(aes(x = sp, y = mean, color = type_sp)) +
    geom_linerange(aes(x = sp, ymin = min, ymax = max, color = type_sp)) +
    facet_wrap(~ var, strip.position = "left", ncol = 1) +
    coord_flip() +
    theme_bw() +
    theme(strip.placement = "outside",
        strip.background = element_blank(),
        strip.text.y = element_text(angle = 0, hjust = 0.5),
        # coller les facettes
        panel.spacing = unit(0, "lines"),
        # lignes de en haut et en bas des facettes en pointillées, que en haut
        # juste un axe tout en bas et tout à gauche
        axis.line = element_line(color = "black"),
        axis.ticks = element_blank(),
        axis.title.y = element_blank()) +
    # axe vertical à 0
    geom_hline(yintercept = 0, linetype = "dashed") +
    # ajouter des lignes entre les facettes
    theme(panel.border = element_rect(color = "grey", fill = NA, size = 0.7)) +
    labs(y = "Posterior distribution", color = "Type of species")
```

```{r}
list_coef = 
    c("Intercept", "Basal area", "Humus pH","Organic matter", "Temperature", "CMI", "Conspecific")
ggplot(all_param %>% filter (var %in% list_coef, type == "nb")) +
    geom_point(aes(x = sp, y = mean, color = var)) +
    geom_linerange(aes(x = sp, ymin = min, ymax = max, color = var)) +
    facet_wrap(~ var, strip.position = "left", ncol = 1) +
    coord_flip() +
    theme_bw() +
    theme(strip.placement = "outside",
        strip.background = element_blank(),
        strip.text.y = element_text(angle = 0, hjust = 0.5),
        # coller les facettes
        panel.spacing = unit(0, "lines"),
        # lignes de en haut et en bas des facettes en pointillées, que en haut
        # juste un axe tout en bas et tout à gauche
        axis.line = element_line(color = "black"),
        axis.ticks = element_blank(),
        axis.title.y = element_blank()) +
    # axe vertical à 0
    geom_hline(yintercept = 0, linetype = "dashed") +
    # ajouter des lignes entre les facettes
    theme(panel.border = element_rect(color = "grey", fill = NA, size = 0.7)) +
    labs(y = "Posterior distribution")
```

# 3. Model output for quadratic parameters

With prediction on data range for this parameters :
- epmatorg : [1:25]
- cmi : [30:100]
- tmean : [-1:3]
- ph : [4:5]

```{r}
# Extract the effect and their 95% confidence interval
quadratic <- quadratic_table() # transform = TRUE will transform the data to the non-linear logit log scale
# Plot the results for each quadratic parameters
quadratic_plot("cmi")
quadratic_plot("tmean")
# save quadratic data as rds
saveRDS(quadratic, "output/quadratic.rds")
```

# 4. Model output for disturbances parameters

```{r, fig.height=6, fig.width = 8, warning = FALSE}
# Extract the effect and their 95% confidence interval for each disturbance
# For presence absence and abundance separatly
disturbance_pa <- disturbance_table("pa", FALSE)
disturbance_nb <- disturbance_table("nb", FALSE)
# Plot the results
disturbance_plot(disturbance_pa, 3)
disturbance_plot(disturbance_nb, 3)
# save disturbance data as rds
saveRDS(disturbance_pa, "output/disturbance_pa.rds")
saveRDS(disturbance_nb, "output/disturbance_nb.rds")
```

Récupérer les distribution entières de tous les paramètres pour chaque espèce pour les perturbations


# 5. Posteriors distribution

## 5.1 For co-variables

```{r}
Coef <- c("nb_epmatorg", "nb_ph",
        "nb_tmean", "nb_tmean2", "nb_cmi", "nb_cmi2",
        "pa_epmatorg",  "pa_ph",
        "pa_tmean", "pa_tmean2", "pa_cmi", "pa_cmi2")
```

```{r}
posteriors <- data.frame()
for (sp in Mysp){
    model <- get(paste0("Model_", sp))
    for (c in Coef){
        df = data.frame(posterior = model$BUGSoutput$sims.list[c][[1]], sp_code = sp, coef = c)
        posteriors <- rbind(posteriors, df)
    }
}

posteriors <- posteriors %>% mutate(part = substr(coef, 1,2))
```

```{r}
ggplot(posteriors %>% filter(part == "pa")) +
    geom_density(aes(x = posterior, fill = sp_code), alpha = 0.5, show.legend = FALSE) +
    facet_wrap(~coef, scales = "free") +
    theme_bw() +
ggplot(posteriors %>% filter(part == "nb")) +
    geom_density(aes(x = posterior, fill = sp_code), alpha = 0.5) +
    facet_wrap(~coef, scales = "free") +
    theme_bw()
```

## 5.2 For disturbances

```{r}
Dist_coef <- c("nb_l", "nb_pl", "nb_lpr", "nb_b",  "nb_o", "pa_l",
    "pa_pl", "pa_lpr", "pa_b", "pa_o")

dist_posteriors <- data.frame()
df <- data.frame()
for (sp in Mysp){
    model <- get(paste0("Model_", sp))
    for (c in Dist_coef){
        df = data.frame(posterior = model$BUGSoutput$sims.list[paste0(c)][[1]][,3], sp_code = sp, coef = c,
            pa_intercept = model$BUGSoutput$sims.list["pa_intercept"][[1]][1],
            nb_intercept = model$BUGSoutput$sims.list["nb_intercept"][[1]][2])
        dist_posteriors <- rbind(dist_posteriors, df)
    }
}

# add type of species temperate, boreal, pioneer
dist_posteriors <- dist_posteriors %>%
    mutate(sp_type = case_when(sp_code %in% c("ACERUB","ACESAC", "BETALL") ~ "Temperate",
                        sp_code %in% c("BETPAP", "POPTRE") ~ "Pioneer",
                        sp_code %in% c("ABIBAL", "PICMAR", "PICGLA") ~ "Boreal"),
        sp_type = factor(sp_type, levels = c("Temperate", "Pioneer", "Boreal")))
dist_posteriors <- dist_posteriors %>%
    mutate(part = ifelse(coef %in% c("pa_l0", "pa_pl0", "pa_lpr0", "pa_b0",  "pa_o0"), "Presence", "Abundance"),
    part = factor(part, levels = c("Presence", "Abundance")),
        perturb = case_when(
            coef %in% c("pa_l0", "nb_l0") ~ "Clearcut",
            coef %in% c("pa_pl0", "nb_pl0") ~ "Partial cut",
            coef %in% c("pa_lpr0", "nb_lpr0") ~ "Logging + protection",
            coef %in% c("pa_b0", "nb_b0") ~ "Fire",
            coef %in% c("pa_o0", "nb_o0") ~ "Outbreak"),
        real_post = ifelse(part == "Presence", inv.logit(pa_intercept + posterior), exp(nb_intercept + posterior)))

```

```{r}
ggplot(dist_posteriors) +# %>% filter(perturb %in% c("Clearcut", "Partial cut"))) +
    geom_density(aes(x = posterior, fill = sp_type), alpha = 0.55) +
    facet_grid(part ~ perturb) +
    scale_fill_manual(name = "Species type", values = c("#2e35f9", "#0cd53e", "#df2222")) +
    theme_bw() +
    theme(legend.position = "bottom",
        plot.margin = unit(c(1,1,1,1), "cm")) +
    xlim(-5,5)
```

# 6. Rhat and neff verification (Rhat ~ 1 and neff > 100)

```{r}
Coef <- c("nb_intercept", "nb_epmatorg", "nb_epmatorg2", "nb_ph", "nb_ph2",
        "nb_tmean", "nb_tmean2", "nb_cmi", "nb_cmi2",
        "nb_sp", "nb_sp2", "nb_year", "nb_ba",
        "nb_l[1]","nb_l[2]", "nb_l[3]",
        "nb_pl[1]", "nb_pl[2]", "nb_pl[3]",
        "nb_lpr[1]", "nb_lpr[2]", "nb_lpr[3]",
        "nb_b[1]", "nb_b[2]", "nb_b[3]",
        "nb_o[1]", "nb_o[2]", "nb_o[3]",
        "pa_intercept", "pa_epmatorg", "pa_epmatorg2", "pa_ph", "pa_ph2",
        "pa_tmean", "pa_tmean2", "pa_cmi", "pa_cmi2",
        "pa_sp", "pa_sp2", "pa_year", "pa_ba",
        "pa_l[1]", "pa_l[2]", "pa_l[3]",
        "pa_pl[1]", "pa_pl[2]", "pa_pl[3]",
        "pa_lpr[1]", "pa_lpr[2]", "pa_lpr[3]",
        "pa_b[1]", "pa_b[2]", "pa_b[3]",
        "pa_o[1]", "pa_o[2]", "pa_o[3]")


# Récupérer les Rhat
Rhat <- sapply(Coef, function(x) {
    data.frame(Model_ACERUB$BUGSoutput$summary)[x, "Rhat"]
})
neff <- sapply(Coef, function(x) {
    data.frame(Model_ACERUB$BUGSoutput$summary)[x, "n.eff"]
})
```

```{r}
Rhat
neff
```